---
title: "Reproducible Research: Peer Assessment 1 9/13/2015"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
unzip("activity.zip")
activitydata3 <- read.csv("activity.csv")
```


## What is mean total number of steps taken per day?
This is what the distribution of steps per day looks like...
```{r}
stepsperday <- aggregate(steps ~ date, data = activitydata3, na.rm = TRUE, FUN = sum)
hist(stepsperday$steps)
```

This is the mean number of steps per day:
```{r}
meanstepsperday <- mean(stepsperday$steps)
meanstepsperday
```

This is the median number of steps per day:
```{r}
medianstepsperday <- median(stepsperday$steps)
medianstepsperday
```


## What is the average daily activity pattern?
This is what the average day looks like in terms of steps per interval.
```{r}
stepsperinterval <- aggregate(steps ~ interval, data = activitydata3, na.rm = TRUE, FUN = mean)
plot(stepsperinterval, type = "l")
```


This is the interval that has the highest average steps:
```{r}
maxtime <- which.max(stepsperinterval$steps)
stepsperinterval[maxtime, 1]
```


## Imputing missing values
This is the number of intervals with missing data for steps:
```{r}
sum(is.na(activitydata3$steps))
```
We are going to replace the missing step data with the median for that interval and create a new dataset with those imputed steps.
```{r}
medianperinterval <- aggregate(steps ~ interval, data = activitydata3, na.rm = TRUE, FUN = median)
impd3 <- merge(activitydata3, medianperinterval, by.x = "interval", by.y = "interval")
impd3$imputedsteps <- impd3$steps.x
impd3$imputedsteps <- ifelse(is.na(impd3$imputedsteps), impd3$steps.y, impd3$steps.x)
activitydataimp <- subset(impd3, select = c("interval", "imputedsteps", "date"))
```

This is a histogram of the total steps per day with the imputed dataset:
```{r}
stepsperdayimp <- aggregate(imputedsteps ~ date, data = activitydataimp, na.rm = TRUE, FUN = sum)
hist(stepsperdayimp$imputedsteps)
```

This is the mean number of steps per day in the imputed data:
```{r}
impmeanday <- mean(stepsperdayimp$imputedsteps)
impmeanday
```

This is the median number of steps per day in the imputed data:
```{r}
impmedianday <- median(stepsperdayimp$imputedsteps)
impmedianday
```

This is the difference between the mean steps per day with the original and the imputed data:
```{r}
impmeanday - meanstepsperday
```

This is the difference between the median steps per day with the original and imputed data:
```{r}
impmedianday - medianstepsperday
```

It appears that the imputation decreases the mean and median.


## Are there differences in activity patterns between weekdays and weekends?
Here are two plots comparing average steps across intervals in a day.
You might notice that during the week that steps start a bit earlier (going to work) but during the weekend steps are a bit higher through the middle of the day (don't have to sit at desk).
```{r}
activitydataimp$date <- as.Date(activitydataimp$date, "%Y-%m-%d")
activitydataimp$weekday <- weekdays(activitydataimp$date)
activitydataimp$weekpart <- ifelse((activitydataimp$weekday == "Saturday" | activitydataimp$weekday == "Sunday"), "weekend", "weekday")
compareweekstepsperinterval <- aggregate(imputedsteps ~ interval + weekpart, data = activitydataimp, FUN = mean)
library(lattice)
xyplot(imputedsteps ~ interval | weekpart, compareweekstepsperinterval, type = "l", layout = c(1,2))
```
